---
title: 'Data Analyst for Finance: The Investor Investment Process'
author: "Dinda_Feb"
date: "2024-01-20"
output: html_document
---

Import The Data
```{r import}
df_event <- read.csv('https://storage.googleapis.com/dqlab-dataset/event.csv', stringsAsFactors = F)
dplyr::glimpse(df_event)
```

Change Data Type
```{r tsmp}
library(lubridate)
df_event$created_at <- ymd_hms(df_event$created_at)
dplyr::glimpse(df_event)
```

Summarise nama_event
```{r summary}
library(dplyr)
df_event %>%
  group_by(nama_event) %>%
  summarise(jumlah_event = n(),
            loan = n_distinct(loan_id),
            investor = n_distinct(investor_id))
```

Upload Loan to Marketplace Table
```{r upload}
df_market_place <- df_event %>%
  filter(nama_event == 'loan_to_marketplace') %>%
  select(loan_id, marketplace=created_at)
head(df_market_place, 5)
```


Investor View Loan Table 
```{r view}
df_view_loan <- df_event %>%
  filter(nama_event == "investor_view_loan") %>%
  group_by(loan_id, investor_id) %>%
  summarise(jumlah_view = n(),
            pertama_view = min(created_at),
            terakhir_view = max(created_at),
            .groups = 'drop')
head(df_view_loan, 5)
```

Event Investor Loan Orders and Payments
```{r order}
library(dplyr)
library(tidyr)
df_order_pay <- df_event %>%
  filter(nama_event %in% c("investor_order_loan", "investor_pay_loan")) %>%
  group_by(loan_id, investor_id, nama_event) %>%
  spread(nama_event, created_at) %>%
  select(loan_id, investor_id, order=investor_order_loan, pay=investor_pay_loan)
head(df_order_pay, 5)
```

Combine The Table
```{r loan}
df_loan_invest <- df_market_place %>%
  left_join(df_view_loan, by = 'loan_id') %>%
  left_join(df_order_pay, by = c("loan_id", "investor_id"))
head(df_loan_invest, 5)
```


Relationship Between Total Views and Orders
```{r invest}
df_loan_invest %>%
  mutate(status_order = ifelse(is.na(order), "not_order", "order")) %>%
  count(jumlah_view, status_order) %>%
  spread(status_order, n, fill = 0) %>%
  mutate(persen_order = scales::percent(order/(order + not_order)))
```


The Time Taken to Order a Loan After Viewing Loan Details
```{r time}
df_loan_invest %>%
  filter(!is.na(order)) %>%
  mutate(lama_order_view = as.numeric(difftime(order, pertama_view, units = "mins"))) %>%
  group_by(jumlah_view) %>%
  summarise(total = n(), min = min(lama_order_view), 
            median = median(lama_order_view), 
            mean = mean(lama_order_view), 
            max = max(lama_order_view)) %>%
  mutate_if(is.numeric, ~round(.,2))
```


Average Time of Loan Ordering Since Uploaded Every Week
```{r average time}
library(ggplot2)
df_lama_order_per_minggu <- df_loan_invest %>%
  filter(!is.na(order)) %>%
  mutate(tanggal = floor_date(marketplace, "week"), 
         lama_order = as.numeric(difftime(order, marketplace, units = "hour"))) %>%
  group_by(tanggal)%>%
  summarise(lama_order = median(lama_order))

ggplot(df_lama_order_per_minggu) + 
  geom_line(aes(x=tanggal, y=lama_order)) + 
  theme_bw()+ 
  labs(title = "Rata-rata lama order pada tahun 2020 lebih lama daripada 2019", 
       x="Tanggal", y="Waktu di marketplace sampai di-pesan(jam)")

```


Did Investors Pay for the Placed Orders
```{r pay}
df_bayar_per_minggu <- df_loan_invest%>%
  filter(!is.na(order)) %>%
  mutate(tanggal = floor_date(marketplace, "week")) %>%
  group_by(tanggal) %>%
  summarise(persen_bayar = mean(!is.na(pay)))

ggplot(df_bayar_per_minggu) +
  geom_line(aes(x = tanggal, y = persen_bayar)) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_bw() +
  labs(title = "Sekitar 95% membayar pesanannya di akhir Mei ada outlier karena lebaran", 
       x = "Tanggal", y = "Pesanan yang dibayar")
```


The Time Taken by Investors to Pay for Their Orders
```{r time orders}
df_lama_bayar_per_minggu <- df_loan_invest %>%
  filter(!is.na(pay)) %>%
  mutate(tanggal = floor_date(marketplace, "week"), 
         lama_bayar = as.numeric(difftime(pay, order, units = "hour"))) %>%
  group_by(tanggal)%>%
  summarise(lama_bayar = median(lama_bayar))

ggplot(df_lama_bayar_per_minggu) + 
  geom_line(aes(x=tanggal, y=lama_bayar)) + 
  theme_bw()+ labs(title = "Waktu pembayarannya cenderung lebih buruk 2x lebih lama dari sebelumnya", 
                   x="Tanggal", y="Waktu di pesanan dibayar(jam)")
```

Trend Investor Register
```{r trend}
library(lubridate)
df_investor_register <- df_event %>%
  filter(nama_event == "investor_register") %>%
  mutate(tanggal = floor_date(created_at, "week")) %>%
  group_by(tanggal) %>%
  summarise(investor = n_distinct(investor_id))

ggplot(df_investor_register) +
  geom_line(aes(x=tanggal, y=investor)) +
  theme_bw() +
  labs(title = "Investor register sempat naik di awal 2020 namun sudah turun lagi",
       x="Tanggal", y="Investor Register")
```

Trend in First-time Investor Investments
```{r first-time}
df_investor_pertama_invest <- df_event %>%
  filter(nama_event == "investor_pay_loan") %>%
  group_by(investor_id) %>%
  summarise(pertama_invest = min(created_at)) %>%
  mutate(tanggal = floor_date(pertama_invest, "week")) %>%
  group_by(tanggal) %>%
  summarise(investor = n_distinct(investor_id))

ggplot(df_investor_pertama_invest) +
  geom_line(aes(x=tanggal, y=investor)) +
  theme_bw() +
  labs(title = "Ada kenaikan trend investor pertama invest, namun turun drastis mulai Maret 2020", 
       y="Investor pertama invest")
```


First Investment Cohort Based on Registration Month
```{r cohort registration}
library(tidyr)
df_register_per_investor <- df_event %>%
  filter(nama_event == "investor_register") %>%
  rename(tanggal_register = "created_at") %>%
  mutate(bulan_register = floor_date(tanggal_register, 'month')) %>%
  select(investor_id, tanggal_register, bulan_register)

df_pertama_invest_per_investor <- df_event %>%
  filter(nama_event == "investor_pay_loan") %>%
  group_by(investor_id) %>%
  summarise(pertama_invest = min(created_at))

df_register_per_investor %>%
  left_join(df_pertama_invest_per_investor, by = "investor_id") %>%
  mutate(lama_invest = as.numeric(difftime(pertama_invest, tanggal_register, units = "day")) %/% 30) %>%
  group_by(bulan_register, lama_invest) %>%
  summarise(investor_per_bulan = n_distinct(investor_id), .groups = "drop") %>%
  group_by(bulan_register) %>%
  mutate(register = sum(investor_per_bulan)) %>%
  filter(!is.na(lama_invest)) %>%
  mutate(invest = sum(investor_per_bulan)) %>%
  mutate(percent_invest = scales::percent(invest/register)) %>%
  mutate(breakdown_persen_invest = scales::percent(investor_per_bulan/invest)) %>%
  select(-investor_per_bulan) %>%
  spread(lama_invest, breakdown_persen_invest)
```


Cohort Retention for Investments
```{r cohort retention}
df_investor_per_investor <- df_event %>%
  filter(nama_event == "investor_pay_loan") %>%
  rename(tanggal_invest = created_at) %>%
  select(investor_id, tanggal_invest)

df_pertama_invest_per_investor %>%
  mutate(bulan_pertama_invest = floor_date(pertama_invest, 'month')) %>%
  inner_join(df_investor_per_investor, by = "investor_id") %>%
  mutate(jarak_invest = as.numeric(difftime(tanggal_invest, pertama_invest, units = "day")) %/% 30) %>%
  group_by(bulan_pertama_invest, jarak_invest) %>%
  summarise(investor_per_bulan = n_distinct(investor_id), .groups = "drop") %>%
  group_by(bulan_pertama_invest) %>%
  mutate(investor = max(investor_per_bulan)) %>%
  mutate(breakdown_persen_invest = scales::percent(investor_per_bulan/investor)) %>%
  select(-investor_per_bulan) %>%
  spread(jarak_invest, breakdown_persen_invest) %>%
  select(-`0`)
```

