---
title: 'Data Analyst for Finance: Branch Performance'
author: "Dinda_Feb"
date: "2024-01-20"
output: html_document
---

Import The Data
```{r df_loan}
df_loan <- read.csv("https://storage.googleapis.com/dqlab-dataset/loan_disbursement.csv", stringsAsFactors = F)
str(df_loan)
```

Total Amount in May 2020
```{r df_loan_mei}
library(dplyr)
df_loan_mei <- df_loan %>%
  filter(tanggal_cair >= '2020-05-01', tanggal_cair <= '2020-05-31') %>%
  group_by(cabang) %>%
  summarise(total_amount = sum(amount))
df_loan_mei
```

Top 5 Highest Total Amount by Branch
```{r highest}
library(scales)
df_loan_mei %>%
  arrange (desc(total_amount)) %>%
  head(5)
```

Top 5 Lowest Total Amount by Branch
```{r lowest}
df_loan_mei %>%
  arrange (desc(total_amount)) %>%
  tail(5)
```

The Age of Each Branch
```{r df_cabang_umur}
df_cabang_umur <- df_loan %>%
  group_by(cabang) %>%
  summarise(pertama_cair = min(tanggal_cair)) %>%
  mutate(umur = as.numeric(as.Date("2020-05-15") - as.Date(pertama_cair)) %/% 30)
df_cabang_umur
```

Combine May Performance with Branch Age
```{r df_loan_mei_umur}
df_loan_mei_umur <- df_cabang_umur %>%
  inner_join(df_loan_mei, by='cabang')
df_loan_mei_umur
```

Age-Performance Relationship in May with ScatterPlot
```{r age performance}
library(ggplot2)
ggplot(df_loan_mei_umur, aes(x = umur, y = total_amount)) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Semakin berumur, performa cabang akan semakin baik",
       x="umur(bulan)",
       y = "Total Amount")
```

Low Performance in Each Branch
```{r df_loan_mei_flag}
df_loan_mei_flag <- df_loan_mei_umur %>%
  group_by(umur) %>%
  mutate(Q1 = quantile(total_amount, 0.25),
         Q3 = quantile(total_amount, 0.75),
         IQR = (Q3-Q1)) %>%
  mutate(flag = ifelse(total_amount < (Q1 - IQR), 'rendah', 'baik'))
df_loan_mei_flag %>%
  filter(flag == 'rendah') %>%
  mutate(umur = as.numeric(umur),
         total_amount = as.numeric(total_amount),
         Q1 = as.numeric(Q1),
         Q3 = as.numeric(Q3),
         IQR = as.numeric(IQR))
```

Scatter Plot
```{r scatter plot}
ggplot(df_loan_mei_flag, aes(x = umur, y = total_amount)) +
  geom_point(aes(color = flag)) +
  scale_color_manual(breaks = c("baik", "rendah"),
                     values = c("blue", "red")) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Ada cabang performa rendah padahal tidak termasuk bottom 5", color = "",
       x = "Umur (bulan)",
       y = "Total Amount")
```

Compare Performance at The Same Age
```{r compare performance}
df_loan_mei_flag %>%
  filter(umur == 3) %>%
  inner_join(df_loan, by = "cabang") %>%
  filter(tanggal_cair >= '2020-05-01', tanggal_cair <= '2020-05-31') %>%
  group_by(cabang, flag) %>%
  summarise(jumlah_hari = n_distinct(tanggal_cair), agen_aktif = n_distinct(agen), total_loan_cair = n_distinct(loan_id), avg_amount = mean(amount), total_amount = sum(amount), .groups = "drop") %>%
  arrange(total_amount) %>%
  mutate(jumlah_hari = as.numeric(jumlah_hari),
         agen_aktif = as.numeric(agen_aktif),
         total_loan_cair = as.numeric(total_loan_cair),
         avg_amount = as.numeric(avg_amount),
         total_amount = as.numeric(total_amount))
```

Low Branch Performance in May
```{r low branch}
df_loan_mei_flag %>%
  filter(umur == 3, flag == "rendah") %>%
  inner_join(df_loan, by = "cabang") %>%
  filter(tanggal_cair >= '2020-05-01', tanggal_cair <= '2020-05-31') %>%
  group_by(cabang, agen) %>%
  summarise(jumlah_hari = n_distinct(tanggal_cair), total_loan_cair = n_distinct(loan_id), avg_amount = mean(amount), total_amount = sum(amount), .groups = "drop") %>%
  arrange(total_amount)
```

Good Branch Performance in May
```{r good_branch}
df_loan %>%
  filter(cabang == "AH") %>%
  filter(tanggal_cair >= '2020-05-01', tanggal_cair <= '2020-05-31') %>%
  group_by(cabang, agen) %>%
  summarise(jumlah_hari = n_distinct(tanggal_cair), total_loan_cair = n_distinct(loan_id), avg_amount = mean(amount), total_amount = sum(amount), .groups = "drop") %>%
  arrange(total_amount)
```
